\documentclass[english,12pt]{article}
\usepackage[]{geometry}
\usepackage{tabularx}
\usepackage{blindtext}
\usepackage{enumerate}
\usepackage{parskip} 
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{inconsolata}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{titlesec}
\usepackage{blindtext}
\usepackage{longtable}
\usepackage{multicol}
\usepackage{varwidth}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[style=numeric]{biblatex} % using biblatex, using apa format!
\addbibresource{sources.bib}
\usepackage[svgnames,table]{xcolor}
\usepackage[tableposition=above]{caption}
\usepackage{pifont}
\usepackage{array,ragged2e}
\newcolumntype{P}[1]{>{\RaggedRight\arraybackslash}p{#1}}
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
\hfuzz=0.64pt % Allow hbox to overflow slightly

\begin{document}
\begin{titlepage}
    \null
    \vspace*{2cm}
    
    \begin{center}
        {\Huge \bfseries CSE 363: Computer Security}\\[1.5cm]
        {\Large \bfseries Final Project Proposal}\\[2cm]
        
        \textbf{Authors:} \\[0.5cm]
        Cody Johnston - \textit{cody.johnston@student.nmt.edu}\\
        Cole Johnson - \textit{cole.johnson@student.nmt.edu}\\
        Colin Grandjean - \textit{colin.grandjean@student.nmt.edu}\\
        John Runyon - \textit{john.runyon@student.nmt.edu}\\
        
        New Mexico Institute of Mining and Technology\\
        Socorro, NM 87801, USA\\[2cm]
        
        {\large Date: March 3, 2025}
    \end{center}
    
    \vfill
    \hrule
    \smallskip
    \centerline{\sc New Mexico Tech}
    \smallskip
    \hrule
\end{titlepage}
\frenchspacing
\pagebreak
\section*{Purpose and Outline of Project}
A botnet (the combination of the words robot and network) is a group of network
enabled malware infected. Bots, the infected devices, are controlled by a central
server or servers, often called the “bot-herder” or the “command and control server”
\cite{Heron:2007}.
A botnet can be used for a wide variety of malicious activities. From DDoS attacks to data
theft, botnets can interfere with computer information systems in a variety of ways
and are worthy of studying and considering mitigation techniques. Our paper will focus
on discussing the threats botnets pose, the mitigation techniques currently implemented,
and later we will implement a basic botnet to demonstrate attacks and mitigation
techniques.
\subsection*{Outline of Project}
Here is what we would tentatively like to cover in each section of our paper:
\begin{enumerate}[\bf (a.)]
    \item Introduction
    \begin{enumerate}[(a.)]
        \item What are botnets?
        \item Examples of botnets
        \item Famous cases of botnet attacks
    \end{enumerate}
    \item Taxonomy of Botnets
    \begin{enumerate}[\bf (a.)]
        \item Architecture / Design
        \item Communication Methods
        \item Purpose
    \end{enumerate}
    \item Attack Vectors
    \begin{enumerate}[(a.)]
        \item Vulnerabilities
        \item Phishing
        \item Worm-Like
        \item External Devices
    \end{enumerate}
    \item Mitigation Techniques
    \begin{enumerate}[(a.)]
        \item Firewalls
        \item Intrusion Detection Systems (IDS)
        \item Challenge and Response Tests
    \end{enumerate}
    \item Experimentation
    \begin{enumerate}
        \item Existing botnet frameworks
        \item Developing a basic botnets
        \item Implementing mitigation techniques (if time permits)
    \end{enumerate}
\end{enumerate}
\subsection*{Taxonomy of Botnets}
Botnets can be constructed in a number of 
ways and fulfill many different purposes for an attacker. 
They can either be centralized, using a client-server model, decentralized, 
using a peer-to-peer model, or a hybrid approach that is a combination of centralized and
decentralized architectures \cite{Antonucci:2022}. 
\subsection*{Attack Vectors}
Botnets exploit various attack vectors to infect devices 
and expand their network. The most common methods that botnets 
use are exploiting vulnerabilities, phishing, worm-like propagation, 
and using external devices \cite{Haber2022}. Botnets exploit vulnerabilities such as 
unpatched software, misconfigurations, and security loopholes 
found within operating systems, applications, or IoT devices \cite{Antonucci:2022}.
\subsection*{Mitigation Techniques}
There any a myriad of different techniques that try to mitigate the damage
caused by botnets, or limit their spread altogether. We would like to focus
on a few different techniques: firewalls, intrusion detection systems \cite{Stefan:2000}, and
challenge response tests.
\subsection*{Experimentation and Results}
We first plan to look into existing 
botnet frameworks to examine their methods and features. 
After research into frameworks, we plan to build a simulated botnet, 
with various bot devices running in containers. 
Using a modularized approach we can change the purpose of the botnet 
dynamically to explore the differences. We plan to compare botnet 
communication and architecture methods by running a set of tasks 
for the different methods. If time permits we plan on testing different 
mitigation techniques.
\subsection*{Conclusion}
Our project will explore botnets in a variety of different contexts. 
We will attempt to approach the subject both from a research-based approach,
creating an overview of botnets as they relate to computer security,
and from an experimental approach by attempting to implement a basic version of
a botnet.
\printbibliography
\end{document}
